R for Clinical Study Reports and Submission

### 2021-11-16 project-folder.Rmd

For R users, you already benefit from a well-defined and consistent folder structure.
That is the [R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf).
Every R package developer is required to follow the same convention to organize
their R functions before the R package can be disseminated through the Comprehensive R Archive Network (CRAN).
As a user, you can easily install and use those R packages after installing from CRAN.
There are many good resources to guide developers on R package development,
for example, the [R Packages book](https://r-pkgs.org/) by Hadley Wickham.

@@

With a formal project, it is also important to provide proper documentation for those help functions.
We used `roxygen2` to document functions.
For example, the header below defines each variable in `fmt_est`.
More details can be found in [Chapter 10 of the R Packages book](https://r-pkgs.org/man.html).

@@

The `roxygen2` documentation will be converted into
standard R documentation format, saved as `.Rd` files in the `man/` folder.
This step is automatically handled by `devtools::document()`.

@@

We have learned a few functions in `devtools` to generate content automatically.
Here is a shortlist of tools that can enhance the workflow.

>>>>

For R users, you already benefit from a well-defined and consistent folder structure.
That is the [R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf).
Every R package developer is required to follow the same convention to organize
their R functions before the R package can be disseminated through the Comprehensive R Archive Network (CRAN).
As a user, you can easily install and use those R packages after downloading from CRAN.
There are many good resources to guide developers on R package development,
such as, the [R Packages book](https://r-pkgs.org/) by Hadley Wickham.

@@

For a clinical trial project, it is also important to provide proper documentation for those help functions.
We use `roxygen2` to document functions.
For example, the header below defines each variable in `fmt_est`.
More details can be found in [Chapter 10 of the R Packages book](https://r-pkgs.org/man.html).

@@

The `roxygen2` documentation will be converted into
standard R documentation format, and saved as `.Rd` files in the `man/` folder.
This step is automatically handled by `devtools::document()`.

@@

We have learned a few functions in `devtools` to generate content automatically.
Here is a list of tools that can enhance the workflow.

### 2021-11-16 submission-environment.Rmd

In this chapter, we will focus on automating the creation of
the R environments with R code to accelerate the dry run testing process,
simplifying the ADRG instructions, and making it easy to recreate
different environment settings with reproducible analysis results.

@@

- Create a project folder under `C:/temp/` with a `.Rproj` file;
- Download R `4.1.1` installer from CRAN, install it into `C:/temp/R/`;
- Not install Rtools (by default, `rtools_version = NULL`);
- Create a `.Rprofile` file under the project folder, set `options(repos)`
  to use the specified `repo` (an MRAN snapshot in this example),
  and instructions to set the R binary path in RStudio;
- Create a `.Renviron` file under the project folder and set the
  library path to be the library of the project-specific R installation.

@@

The helper functions `version_*()` and `snapshot_*()` could assist you
in determining specific versions of R and Rtools that are currently available,
besides generating and verifying the snapshot repo links.

@@

If using `cleanslate`, remember to update the ADRG instructions for
executing the analysis programs in R. Mostly, this would simplify
the first three steps on creating a project, installing a specific
version of R, and configuring the package repo location.
For example:

@@

The addin provides a wizard-like interface to help create the environment
with the most important options, although with less flexibility compared
to the functional API demonstrated above.

>>>>

In this chapter, we focus on automating the creation of
the R environments with R code to accelerate the dry run testing process,
simplifying the ADRG instructions, and making it easy to recreate
different environment settings with reproducible analysis results.

@@

- Create a project folder under `C:/temp/` with a `.Rproj` file;
- Download R `4.1.1` installer from CRAN, and install it into `C:/temp/R/`;
- Not install Rtools (by default, `rtools_version = NULL`);
- Create a `.Rprofile` file under the project folder, set `options(repos)`
  to use the specified `repo` (an MRAN snapshot in this example),
  and give instruction to set the R binary path in RStudio;
- Create a `.Renviron` file under the project folder and set the
  library path to be the library of the project-specific R installation.

@@

The helper functions `version_*()` and `snapshot_*()` can assist you
in determining specific versions of R and Rtools that are currently available,
besides generating and verifying the snapshot repo links.

@@

If you use `cleanslate`, remember to update the ADRG instructions for
executing the analysis programs in R. Mostly, this can simplify
the first three steps on creating a project, installing a specific
version of R, and configuring the package repo location.
For example:

@@

The addin provides a wizard-like interface to help create the environment
with the most important options, yet with less flexibility compared
to the functional API demonstrated above.
